---
title: "TaASN-B2 paper figures"
author: "Joseph Oddy"
date: "28/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rstatix)
library(ggpubr)
library(RColorBrewer)
library(dplyr)
library(waffle)
library(plyr)
theme_set(theme_pubclean())
```

## Waffle plot

Import data (pie) as csv and create count dataframe of TaASN-B2 positive and negative varieties grouped by milling type (as classified by UK Flour millers).

```{r, echo = FALSE}
pie<-read.csv("C:/Users/Joe/OneDrive - Rothamsted Research/PhD/ASN2 B project/R data and scripts/201218 pie chart data.csv")
```

```{r}
pie %>% 
  dplyr::group_by(Type) %>% 
  dplyr::count(Type, ASN_B2) -> counts_df
```

Plot this as waffle plot with Github version of waffle package (includes geom_waffle function).
```{r}

allwaffle <- ggplot(counts_df, aes(fill = ASN_B2, values = n)) +
  geom_waffle(color = "white", size = 1, n_rows = 5, flip = TRUE) +
  facet_wrap(~Type, nrow = 1, strip.position = "bottom") +
  scale_x_discrete()   + 
  scale_y_continuous(labels = function(x) x * 5, expand = c(0,0)) + coord_equal() +
  labs(x = "NABIM groups", y = "Variety count") +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.ticks.y = element_line(),legend.position="top") +
  guides(fill = guide_legend(reverse = TRUE))
allwaffle
```

## qPCR plots

Import data (qpcr), subset by variety, and calculate standard errors for each variety/timepoint/homeologue combination.

```{r, echo = FALSE, message=FALSE}
qpcr<-read.csv("C:/Users/Joe/OneDrive - Rothamsted Research/PhD/ASN2 B project/ASN2 B paper ea/qPCR/R scripts and data/201104 summary data restructured.csv", stringsAsFactors = TRUE)

cad <- subset(qpcr, 誰..Genotype=="Cadenza")
cla <- subset(qpcr, 誰..Genotype=="Claire")
dux <- subset(qpcr, 誰..Genotype=="Duxford")
spa <- subset(qpcr, 誰..Genotype=="Spark")
```

```{r, message=FALSE}
cad.summary.se <- cad %>% # the names of the new data frame and the data frame to be summarised
  group_by(Timepoint, Homeologue) %>%   # the grouping variable
  dplyr::summarise( # calculates the mean of each group
    sd_expression = sd(Expression),
    SE_expression = sd(Expression)/sqrt(n()),
    Expression = mean(Expression),# calculates the standard deviation of each group
    n_expression = n()) %>%
  mutate(se.expression = sd_expression / sqrt(n_expression),
         lower.ci.expression = Expression - qt(1 - (0.05 / 2), n_expression - 1) * se.expression,
         upper.ci.expression = Expression + qt(1 - (0.05 / 2), n_expression - 1) * se.expression) # calculates the standard error of each group

cla.summary.se <- cla %>% # the names of the new data frame and the data frame to be summarised
  group_by(Timepoint, Homeologue) %>%   # the grouping variable
  dplyr::summarise( # calculates the mean of each group
    sd_expression = sd(Expression),
    SE_expression = sd(Expression)/sqrt(n()),
    Expression = mean(Expression),# calculates the standard deviation of each group
    n_expression = n()) %>%
  mutate(se.expression = sd_expression / sqrt(n_expression),
         lower.ci.expression = Expression - qt(1 - (0.05 / 2), n_expression - 1) * se.expression,
         upper.ci.expression = Expression + qt(1 - (0.05 / 2), n_expression - 1) * se.expression) # calculates the standard error of each group

dux.summary.se <- dux %>% # the names of the new data frame and the data frame to be summarised
  group_by(Timepoint, Homeologue) %>%   # the grouping variable
  dplyr::summarise( # calculates the mean of each group
    sd_expression = sd(Expression),
    SE_expression = sd(Expression)/sqrt(n()),
    Expression = mean(Expression),# calculates the standard deviation of each group
    n_expression = n()) %>%
  mutate(se.expression = sd_expression / sqrt(n_expression),
         lower.ci.expression = Expression - qt(1 - (0.05 / 2), n_expression - 1) * se.expression,
         upper.ci.expression = Expression + qt(1 - (0.05 / 2), n_expression - 1) * se.expression) # calculates the standard error of each group

spa.summary.se <- spa %>% # the names of the new data frame and the data frame to be summarised
  group_by(Timepoint, Homeologue) %>%   # the grouping variable
  dplyr::summarise( # calculates the mean of each group
    sd_expression = sd(Expression),
    SE_expression = sd(Expression)/sqrt(n()),
    Expression = mean(Expression),# calculates the standard deviation of each group
    n_expression = n()) %>%
  mutate(se.expression = sd_expression / sqrt(n_expression),
         lower.ci.expression = Expression - qt(1 - (0.05 / 2), n_expression - 1) * se.expression,
         upper.ci.expression = Expression + qt(1 - (0.05 / 2), n_expression - 1) * se.expression) # calculates the standard error of each group
```

Plot qPCR data with calculated standard errors.

```{r, fig.height = 7.5, fig.width = 7.5, fig.align = "center"}
cadplot <- ggplot(cad, aes(Timepoint, Expression, color = Homeologue)) +
  geom_col(data = cad.summary.se, position = position_dodge(0.8), 
           width = 0.7, fill = "white") +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), size = 1) +
  geom_errorbar(aes(ymin = Expression-SE_expression, ymax = Expression+SE_expression), data = cad.summary.se, 
                width = 0.2, position = position_dodge(0.8)) +
  ylim(0,0.8) + ylab(expression(RE~(E^"-Ct target"/E^"-Ct reference"))) + 
  xlab("Timepoint")+ theme_bw()

claplot <- ggplot(cla, aes(Timepoint, Expression, color = Homeologue)) +
  geom_col(data = cla.summary.se, position = position_dodge(0.8), 
           width = 0.7, fill = "white") +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), size = 1) +
  geom_errorbar(aes(ymin = Expression-SE_expression, ymax = Expression+SE_expression), data = cla.summary.se, 
                width = 0.2, position = position_dodge(0.8))+
  ylim(0,0.4)+ ylab(expression(RE~(E^"-Ct target"/E^"-Ct reference"))) + 
  xlab("Timepoint")+ theme_bw()

duxplot <- ggplot(dux, aes(Timepoint, Expression, color = Homeologue)) +
  geom_col(data = dux.summary.se, position = position_dodge(0.8), 
           width = 0.7, fill = "white") +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), size = 1) +
  geom_errorbar(aes(ymin = Expression-SE_expression, ymax = Expression+SE_expression), data = dux.summary.se, 
                width = 0.2, position = position_dodge(0.8)) +
  ylim(0,2.0) + ylab(expression(RE~(E^"-Ct target"/E^"-Ct reference"))) + 
  xlab("Timepoint") + theme_bw()

spaplot <- ggplot(spa, aes(Timepoint, Expression, color = Homeologue)) +
  geom_col(data = spa.summary.se, position = position_dodge(0.8), 
           width = 0.7, fill = "white") +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), size = 1) +
  geom_errorbar(aes(ymin = Expression-SE_expression, ymax = Expression+SE_expression), data = spa.summary.se, 
                width = 0.2, position = position_dodge(0.8)) +
  ylim(0,2.5) + ylab(expression(RE~(E^"-Ct target"/E^"-Ct reference"))) + 
  xlab("Timepoint") + theme_bw()

ggarrange(cadplot, duxplot, claplot, spaplot, labels = c("a.","b.","c.","d."), 
          common.legend = TRUE, ncol = 2, nrow = 2)
```

## Box-and-scatter plots

Data for each plot (combined, first (trial), second (trial), REML) are first imported as separate csv files.

```{r, echo = FALSE}
combined <- read.csv("C:/Users/Joe/OneDrive - Rothamsted Research/PhD/ASN2 B project/R data and scripts/210408 combined years data.csv")

combined$ASN_B2 <- factor(combined$ASN_B2, levels=c("Present", "Absent"))

second <- read.csv("C:/Users/Joe/OneDrive - Rothamsted Research/PhD/ASN2 B project/R data and scripts/210330 second year data.csv")

first <- read.csv("C:/Users/Joe/OneDrive - Rothamsted Research/PhD/ASN2 B project/R data and scripts/210330 first year data.csv")

first$ASN_B2 <- factor(first$ASN_B2, levels=c(" Present", " Absent"))

reml <- read.csv("C:/Users/Joe/OneDrive - Rothamsted Research/PhD/ASN2 B project/R data and scripts/201216 ASN2 both years REML predicted means.csv")

reml$ASN_B2 <- factor(reml$ASN_B2, levels=c("Present", "Absent"))
```

First plot (A.) looks at each trial without separating the effect of sulphur:

```{r}
yearplot <- ggplot(combined, aes(x = ASN_B2, y = Asn))

yearplot <- yearplot + geom_boxplot(outlier.shape = NA) + 
  geom_jitter(aes(colour = factor(Treatment)),width = 0.2, size = 1) +
  ylab(expression(Free~asparagine~(mmol~per~kg))) + xlab("Sulphur Treatment (2011 - 2012)") +
  stat_summary(aes(x=ASN_B2, y=Asn, group = Season),fun=mean, geom="line", linetype = "dashed")+ 
  stat_summary(aes(x=ASN_B2, y=Asn, group = Season),fun=mean, geom="point", shape = 3)+
  theme(legend.key=element_blank(),legend.background=element_blank())+ theme_bw() +
  facet_wrap(. ~ Season, scales = "free_y") +
  scale_colour_manual(values = c("#004D40", "#FFC107"))+
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.3)))

```

Second and third plots (B. and C.) separate the effects of sulphur:

```{r}
sbp1 <- ggplot(first, aes(x = ASN_B2, y = Asn))

sbp1 <- sbp1 + geom_boxplot(outlier.shape = NA) + 
  geom_jitter(aes(colour = factor(Treatment)),width = 0.2, size = 1) +
  ylab(expression(Free~asparagine~(mmol~per~kg))) + xlab("Sulphur Treatment (2011 - 2012)") +
  stat_summary(aes(x=ASN_B2, y=Asn, group = Treatment),fun=mean, geom="line", linetype = "dashed")+ 
  stat_summary(aes(x=ASN_B2, y=Asn, group = Treatment),fun=mean, geom="point", shape = 3)+
  theme(legend.key=element_blank(),legend.background=element_blank())+ theme_bw() +
  facet_wrap(. ~ Treatment, scales = "free_y") +
  scale_colour_manual(values = c("#004D40", "#FFC107"))+
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.3)))

sbp2 <- ggplot(second, aes(x = ASN_B2, y = Asn))

sbp2 <- sbp2 + geom_boxplot(outlier.shape = NA) + 
  geom_jitter(aes(colour = factor(Treatment)),width = 0.2, size = 1) +
  ylab(expression(Free~asparagine~(mmol~per~kg))) + xlab("Sulphur Treatment (2012 - 2013)") +
  stat_summary(aes(x=ASN_B2, y=Asn, group = Treatment),fun=mean, geom="line", linetype = "dashed")+ 
  stat_summary(aes(x=ASN_B2, y=Asn, group = Treatment),fun=mean, geom="point", shape = 3)+
  theme(legend.key=element_blank(),legend.background=element_blank())+ theme_bw() +
  facet_wrap(. ~ Treatment, scales = "free_y")+
  scale_colour_manual(values = c("#004D40", "#FFC107"))+
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.3)))
```

Fourth plot (D.) uses predicted means from GenStat REML analysis for those varieties present in both trials:

```{r}
remlplot <- ggplot(reml, aes(x = ASN_B2, y = Asn))

remlplot <- remlplot + geom_boxplot(outlier.shape = NA) + 
  geom_jitter(aes(colour = factor(Treatment)),width = 0.2, size = 1) +
  ylab(expression(Free~asparagine~(mmol~per~kg))) + xlab("ASN-B2 (REML)") +
  stat_summary(aes(x=ASN_B2, y=Asn, group = Treatment),fun=mean, geom="line", linetype = "dashed")+ 
  stat_summary(aes(x=ASN_B2, y=Asn, group = Treatment),fun=mean, geom="point", shape = 3)+
  theme(legend.key=element_blank(),legend.background=element_blank())+ theme_bw() +
  facet_wrap(. ~ Treatment, scales = "free_y") +
  scale_colour_manual(values = c("#004D40", "#FFC107"))+
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.3)))
```

All box-and-scatter plots plotted together:

```{r, fig.height = 7.5, fig.width = 12.5, fig.align = "center"}
all <- ggarrange(yearplot, sbp1, sbp2, remlplot, labels = c("a.","b.","c.", "d."), 
                 common.legend = TRUE, ncol = 4, nrow = 1)

all
```

